\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=1in]{geometry}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\title{Strict Positivity of the Dirichlet Square Root Behind OEIS A317940}
\author{Dominic Dabish\\\small AI-assisted discovery and Lean formalization; independent review pending}
\date{July 2026}
\begin{document}
\maketitle
\begin{abstract}
Let $g$ be the multiplicative arithmetic function determined by
$g(p^e)=2^{A005187(e)}$, where
$A005187(e)=\sum_{k\ge0}\lfloor e/2^k\rfloor$.
If $f(1)=1$ and $f*f=g$ under Dirichlet convolution, OEIS A317940 records the numerators of $f(n)$ and asks whether they are nonnegative. We prove the stronger statement $f(n)>0$ for every $n\ge1$. The proof reduces the problem to positivity of the formal square root of a binary Euler product. The exact Formal Conjectures statement has also been verified in Lean 4.
\end{abstract}

\section{Local reduction}
Because $g$ is multiplicative, its normalized Dirichlet square root is multiplicative. Put $c_e=f(p^e)$. Then
\begin{equation}\label{local}
\sum_{j=0}^{e}c_jc_{e-j}=2^{A005187(e)}.
\end{equation}
The recurrence $A005187(e)=e+A005187(\lfloor e/2\rfloor)$ yields
$A005187(e)=2e-s_2(e)$, where $s_2(e)$ is the binary digit sum. Set $b_e=2^{-s_2(e)}$. Then
$b_{2r}=b_r$ and $b_{2r+1}=b_r/2$.

\section{The binary Euler product}
Let $B(z)=\sum_{e\ge0}b_ez^e$. Unique binary expansion gives
\[
B(z)=\prod_{r\ge0}\left(1+\frac{z^{2^r}}2\right).
\]
Let $A(z)^2=B(z)$ with $A(0)=1$, and write $A(z)=\sum a_ez^e$.

\begin{lemma}
Every nonconstant coefficient of $\log A(z)$ is strictly positive.
\end{lemma}
\begin{proof}
Write $n=2^vm$ with $m$ odd. Expanding the formal logarithm gives
\[
[z^n]\log A(z)=\frac1{2m}\left(2^{-m}-\sum_{j=1}^{v}2^{-j-2^jm}\right).
\]
The negative sum is strictly smaller than
$2^{-2m}\sum_{j\ge1}2^{-j}=2^{-2m}<2^{-m}$.
\end{proof}

\begin{theorem}
Every coefficient $a_e$ is strictly positive.
\end{theorem}
\begin{proof}
Since $A=\exp(\log A)$ and every nonconstant coefficient of $\log A$ is positive, the coefficient of $z^e$ in $A$ has the positive linear contribution $[z^e]\log A$, while all higher exponential contributions are nonnegative.
\end{proof}

\section{Return to Dirichlet convolution}
Set $c_e=4^ea_e$. Rescaling $A(z)^2=B(z)$ by $z\mapsto4z$ gives
\[
\sum_{j=0}^{e}c_jc_{e-j}=4^eb_e=2^{A005187(e)},
\]
so \eqref{local} holds and $c_e>0$.
Define the multiplicative arithmetic function $h$ by $h(p^e)=c_e$.
Then $h*h=g$. Separating the endpoint divisors in this equality gives, for $n>1$,
\[
h(n)=\frac12\left(g(n)-\sum_{\substack{d\mid n\\1<d<n}}h(d)h(n/d)\right).
\]
This is exactly the recursion defining the rational function behind A317940. By strong induction, $f=h$, and hence $f(n)>0$ for all $n\ge1$.

\begin{theorem}
The rational function underlying OEIS A317940 is strictly positive at every positive integer. In particular, A317940 has no negative or zero signed numerator.
\end{theorem}

\section{Generalization}
For $0<q\le1$ and $\alpha>0$, every coefficient of
\[
\prod_{r\ge0}(1+qz^{2^r})^\alpha
\]
is strictly positive. Indeed, for $n=2^vm$ with $m$ odd,
\[
[z^n]\log\prod_{r\ge0}(1+qz^{2^r})^\alpha
=\frac{\alpha}{m}\left(q^m-\sum_{j=1}^{v}\frac{q^{2^jm}}{2^j}\right)>0.
\]

\section*{Formal verification}
A Lean 4 proof checks the exact theorem named \texttt{A317940\_f\_nonnegative} against the Google DeepMind Formal Conjectures definitions. The proof contains no placeholders or custom axioms. A complete upstream-repository build and axiom audit accompany the source.
\end{document}
