\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\hypersetup{
  pdftitle={The 2n-4 Bound for Checkerboard No-Three-in-Line Sets},
  pdfauthor={Dominic A. Dabish},
  pdfsubject={Checkerboard no-three-in-line sets; quadratic line-cover certificates; Lean 4 formalization}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\Dmono}{D_{\mathrm{mono}}}
\newcommand{\Ceps}{C_{\varepsilon}}

\title{The $2n-4$ Bound for Checkerboard No-Three-in-Line Sets}
\author{Dominic A. Dabish\\\small Independent researcher, San Diego, California, USA}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
Let $\Dmono(n)$ be the largest number of points that can be chosen from one parity class of the $n\times n$ integer grid with no three collinear. Prellberg proved the elementary bound $\Dmono(n)\le 2n-2$ and proposed the sharper inequality $\Dmono(n)\le 2n-4$ for $n\ge 6$ as a boundary-forcing heuristic. We prove this inequality for every $n\ge 6$.

The proof is an explicit four-direction linear-programming dual certificate. On every sufficiently large odd or even board, quadratic weights on rows, columns, and the two diagonal families give constant coverage of every point in the chosen parity class, while their exact total cost lies strictly below the threshold for $2n-3$ selected points. The thin colour of the $7\times7$ board is handled by a small integer certificate, and the two $6\times6$ colour classes by a finite kernel-checked Boolean verification. A complete Lean~4 formalization proves the exact all-$n$ theorem and audits its axioms.
\end{abstract}

\section{Introduction}

Write
\[
G_n=\{0,1,\ldots,n-1\}^2
\]
and, for $\varepsilon\in\{0,1\}$, let
\[
C_\varepsilon(n)=\{(x,y)\in G_n:x+y\equiv\varepsilon\pmod 2\}.
\]
A set is \emph{no-three-in-line} (NTIL) if it contains no three collinear points. Define
\[
\Dmono(n,\varepsilon)=\max\{|S|:S\subseteq C_\varepsilon(n)\text{ is NTIL}\},
\qquad
\Dmono(n)=\max_{\varepsilon\in\{0,1\}}\Dmono(n,\varepsilon).
\]

Prellberg introduced this fixed-parity variant and observed that the two monochromatic diagonal families of slopes $\pm1$ give
\[
\Dmono(n)\le 2n-2.
\]
He further recorded a boundary-forcing mechanism suggesting, but not proving,
\[
\Dmono(n)\le 2n-4\qquad(n\ge6).
\]
The purpose of this note is to prove that inequality exactly.

The proof uses only a necessary consequence of the full NTIL condition: every row, column, slope-$1$ diagonal, and slope-$(-1)$ diagonal contains at most two selected points. Thus a certificate using these four line families gives an upper bound for the original all-slope problem, not merely for a redefined four-direction problem.

\begin{theorem}[Main theorem]\label{thm:main}
For every integer $n\ge6$ and each $\varepsilon\in\{0,1\}$,
\[
\Dmono(n,\varepsilon)\le 2n-4.
\]
Consequently $\Dmono(n)\le2n-4$ for every $n\ge6$.
\end{theorem}

\section{A weighted line-cover lemma}

Let $\mathcal L$ consist of all rows, columns, and diagonals of slopes $\pm1$ that meet $G_n$. Assign a nonnegative weight $w_\ell$ to every $\ell\in\mathcal L$. For $p\in G_n$, let
\[
\operatorname{cov}(p)=\sum_{\ell\ni p}w_\ell,
\qquad
\operatorname{cost}(w)=2\sum_{\ell\in\mathcal L}w_\ell.
\]
The factor $2$ is the capacity of each line in an NTIL set.

\begin{lemma}[Line-cover certificate]\label{lem:certificate}
Suppose $S\subseteq G_n$ is NTIL and every $p\in S$ satisfies $\operatorname{cov}(p)\ge q$, where $q>0$. Then
\[
q|S|\le \operatorname{cost}(w).
\]
In particular, if $\operatorname{cost}(w)<q(k+1)$, then $|S|\le k$.
\end{lemma}

\begin{proof}
Double-count weighted incidences:
\[
q|S|
\le \sum_{p\in S}\operatorname{cov}(p)
=\sum_{\ell\in\mathcal L}w_\ell|S\cap\ell|
\le2\sum_{\ell\in\mathcal L}w_\ell.
\]
The last assertion follows because $|S|$ is an integer.
\end{proof}

This is the finite dual-certificate interpretation of the four-direction relaxation used by Prellberg.

\section{Odd side length}

Let $n=2m+1$. The parity class containing $(0,0)$ is the \emph{fat} class and the other is the \emph{thin} class. Put $N=2m$.

For every row and every column with index $i\in\{0,\ldots,2m\}$, assign weight
\[
R_i=(2i-2m)^2.
\]
For a diagonal indexed by $j\in\{0,\ldots,4m\}$, assign
\[
B_j=2\bigl((2m)^2-(j-2m)^2\bigr)
\]
when $j$ has the parity of the chosen colour, and weight $0$ otherwise. Here the two diagonal indices are $j=x+y$ and $j=x-y+2m$.

The weights are nonnegative. If $(x,y)$ lies in the chosen parity class, both incident diagonal indices have the permitted parity, and
\begin{align*}
&(2x-2m)^2+(2y-2m)^2\\
&\quad+2\bigl((2m)^2-(x+y-2m)^2\bigr)
+2\bigl((2m)^2-(x-y)^2\bigr)
=16m^2.
\end{align*}
Thus every admissible point has constant coverage
\[
q=16m^2.
\]

The row or column sum is
\begin{equation}\label{eq:odd-axis}
\sum_{i=0}^{2m}(2i-2m)^2
=\frac{4m(m+1)(2m+1)}3.
\end{equation}
For even diagonal indices, corresponding to the fat class,
\begin{equation}\label{eq:odd-fat-diag}
\sum_{\substack{0\le j\le4m\\j\equiv0\ (2)}}B_j
=\frac{8m(2m-1)(2m+1)}3.
\end{equation}
For odd diagonal indices, corresponding to the thin class,
\begin{equation}\label{eq:odd-thin-diag}
\sum_{\substack{0\le j\le4m\\j\equiv1\ (2)}}B_j
=\frac{4m(8m^2+1)}3.
\end{equation}
These are elementary sums of squares.

\subsection{The fat class}

Using two axis families and two diagonal families, the certificate cost is
\[
C_{\mathrm{fat}}
=\frac{16m(10m^2+3m-1)}3.
\]
The target bound is $|S|\le4m-2$. We compute
\begin{align*}
q(4m-1)-C_{\mathrm{fat}}
&=\frac{16m}{3}(2m^2-6m+1)>0
\end{align*}
for every $m\ge3$. Lemma~\ref{lem:certificate} therefore gives
\[
|S|\le4m-2=2n-4
\]
for every odd fat board with $n\ge7$.

\subsection{The thin class}

The certificate cost is now
\[
C_{\mathrm{thin}}
=\frac{16m(10m^2+3m+2)}3.
\]
For $m\ge4$,
\[
q(4m-1)-C_{\mathrm{thin}}
=\frac{16m}{3}(2m^2-6m-2)>0.
\]
Hence $|S|\le4m-2=2n-4$ for every odd thin board with $n\ge9$.

The remaining thin board is $n=7$. Give weight $1$ to each boundary row or column with index $0$ or $6$. For each of the two diagonal families, indexed by $j\in\{0,\ldots,12\}$, define
\[
f(j)=
\begin{cases}
2,&|j-6|=1,\\
1,&|j-6|=3,\\
0,&\text{otherwise}.
\end{cases}
\]
Only odd $j$ occur in the thin class. The total line-weight sum is $16$, so the capacity-two cost is $32$. Up to the dihedral symmetries of the square, the thin class has four point orbits; their coverages are shown below.

\begin{center}
\begin{tabular}{ccc}
\toprule
Representative & Orbit size & Coverage\\
\midrule
$(0,1)$ & $8$ & $3$\\
$(0,3)$ & $4$ & $3$\\
$(1,2)$ & $8$ & $3$\\
$(2,3)$ & $4$ & $4$\\
\bottomrule
\end{tabular}
\end{center}
Thus $q=3$ and $32<3\cdot11$, proving $|S|\le10=2\cdot7-4$.

\section{Even side length}

Let $n=2m$ and put $N=2m-1$. For each row and column index $i\in\{0,\ldots,2m-1\}$ assign
\[
R_i=(2i-N)^2.
\]
For a diagonal index $j\in\{0,\ldots,2N\}$ assign
\[
B_j=2\bigl(N^2-(j-N)^2\bigr).
\]
On the sum diagonals $j=x+y$, retain the parity of the chosen colour. On the shifted difference diagonals $j=x-y+N$, retain the opposite parity, since $N$ is odd.

Every point of the chosen colour then has coverage
\begin{align*}
&(2x-N)^2+(2y-N)^2\\
&\quad+2\bigl(N^2-(x+y-N)^2\bigr)
+2\bigl(N^2-(x-y)^2\bigr)
=4N^2.
\end{align*}
Hence $q=4N^2$. The axis sum is
\[
\sum_{i=0}^{2m-1}(2i-N)^2
=\frac{2m(4m^2-1)}3.
\]
The two selected diagonal parity classes together partition all diagonal indices, and
\[
\sum_{j=0}^{2N}B_j
=\frac{2N(4m-3)(4m-1)}3.
\]
The resulting cost is
\[
C_{\mathrm{even}}
=\frac{4N(20m^2-14m+3)}3.
\]
For $m\ge4$,
\begin{align*}
q(4m-3)-C_{\mathrm{even}}
&=\frac{8(2m-1)}3(2m^2-8m+3)>0.
\end{align*}
Therefore every even board with $n\ge8$ satisfies
\[
|S|\le4m-4=2n-4.
\]

\section{The \texorpdfstring{$6\times6$}{6x6} base case}

Each parity class of the $6\times6$ board contains $18$ points. The quadratic certificate above has objective exactly $9$ and therefore does not by itself exclude nine points. The remaining assertion is finite:

\begin{proposition}\label{prop:n6}
Every subset of either parity class of the $6\times6$ board that meets every row, column, and slope-$\pm1$ diagonal in at most two points has size at most $8$.
\end{proposition}

The proposition is verified in two independent exact forms in the accompanying package. A standard-library Python checker enumerates the $\binom{18}{9}=48{,}620$ nine-point subsets in each colour and rejects each one by a principal-line capacity violation. The Lean development represents an arbitrary selection by $18$ Boolean variables and proves the same finite statement with the kernel-checked bit-vector decision procedure \texttt{bv\_decide}. Since every all-slope NTIL set obeys these four-direction capacities, this proves the required $n=6$ case.

\begin{proof}[Proof of Theorem~\ref{thm:main}]
The odd fat case holds for $n\ge7$, the odd thin case for $n\ge9$ with a separate certificate at $n=7$, and the even case for $n\ge8$. Proposition~\ref{prop:n6} handles $n=6$.
\end{proof}

\section{Formal verification and reproducibility}

The complete theorem is formalized in Lean~4 with Mathlib. The public declaration is
\begin{verbatim}
theorem Checkerboard.checkerboard_upper_all_n {n parity : Nat}
    (hn : 6 <= n) (hp : Or (parity = 0) (parity = 1))
    (s : Finset (Point n))
    (hcolor : Monochromatic parity s)
    (hntil : NoThreeInLine s) :
    s.card <= 2 * n - 4
\end{verbatim}
Here \texttt{NoThreeInLine} quantifies over every triple of distinct selected points and uses the exact integer determinant, so the formal statement is the genuine all-slope theorem. The four-direction constraints are derived as lemmas from that hypothesis.

The formal dependency chain includes:
\begin{enumerate}[leftmargin=*]
\item the finite point model, checkerboard parity classes, and integer determinant;
\item derivation of the four principal line capacities from full NTIL;
\item the rational weighted line-cover theorem;
\item exact polynomial sum identities, nonnegativity, point coverage, and objective formulas for the odd and even quadratic certificates;
\item the exceptional $7\times7$ thin certificate;
\item the two finite $6\times6$ Boolean base cases;
\item the all-$n$ parity assembly and an explicit axiom audit.
\end{enumerate}

The project is pinned to Lean~4.32.0. The build workflow rejects \texttt{sorry}, \texttt{admit}, project-defined axioms, \texttt{native\_decide}, opaque shortcuts, and any occurrence of \texttt{sorryAx} in the compiled audit output. Exact source commit and workflow identifiers are recorded in the accompanying reproducibility document.

\section{Scope and open questions}

Theorem~\ref{thm:main} proves the finite upper bound proposed in Prellberg's paper. It does not determine $\Dmono(n)$ exactly for every $n$, and it does not settle the conjectured asymptotic value
\[
\lim_{n\to\infty}\frac{\Dmono(n)}n.
\]
In particular, proving a matching all-slope lower bound requires controlling every rational slope and is separate from the four-direction dual argument used here.

\paragraph{AI assistance disclosure.}
OpenAI's ChatGPT was used to assist with proof exploration, Lean formalization, code generation, and manuscript editing. The mathematical statement is independently represented in a kernel-checked Lean proof, and the finite base cases have separate exact checkers. Responsibility for the manuscript and its claims rests with the named author.

\begin{thebibliography}{9}

\bibitem{prellberg-checkerboard}
T. Prellberg,
\newblock No-three-in-line sets on the checkerboard grid,
\newblock \emph{arXiv preprint} arXiv:2605.09215, 2026.

\bibitem{schrijver}
A. Schrijver,
\newblock \emph{Theory of Linear and Integer Programming},
\newblock Wiley, 1986.

\bibitem{lean4}
L. de Moura and S. Ullrich,
\newblock The Lean 4 theorem prover and programming language,
\newblock in \emph{Automated Deduction -- CADE 28}, Lecture Notes in Computer Science 12699, Springer, 2021, pp. 625--635.

\bibitem{mathlib}
The Mathlib Community,
\newblock The Lean mathematical library,
\newblock in \emph{Proceedings of the 9th ACM SIGPLAN International Conference on Certified Programs and Proofs}, 2020, pp. 367--381.

\end{thebibliography}

\end{document}
